FlowFPX: Nimble Tools for Debugging Floating-Point Exceptions

### 2023-08-15 paper.tex
Finish first column of local feedback

\begin{abstract}
  Reliable numerical computations are central to high-performance computing and machine learning.
  We present FlowFPX: a tool set for debugging floating-point code.
  FlowFPX is composed of two primary parts: FloatTracker, a Julia-based tool for tracking the onset and flow of floating-point exceptions; and CSTG, a stack trace visualizer.
  FlowTracker's design exploits Julia's operator overloading to trace exception flows and even inject exceptions to accelerate testing.
  We present intuitive visualizations of summarized exception flows including how they are generated, propagated and killed, thus helping with debugging and repair.
\end{abstract}

@@

Reliable numeric computations are central to high-performance computing,
machine learning, and scientific applications.
Yet, the \fp{} arithmetic that powers these computations is fundamentally
unreliable~(\cref{s:background}).
Exceptional values, namely Not a Number (\Nan{}) and infinity (\Inf{}),
can and often do arise thanks to culprits such as roundoff error,
catastrophic cancellation, singular matrices, and vanishing
derivatives~\cite{sdjmrstp-pc-2022,ddghlllprr-correctness-2022,gllprt-correctness-2021,fpchecker-reports,llg-soap-2022,bllmg-xloop-2022}.
When exceptions happen, domain experts must take time away from their main
focus and trek down a rabbit hole to find the source of the error.
There is little tool support to assist in this difficult task,
thus (unsurpringly) a quick GitHub search reports over 4,000 open issues
related to \NaN{} or \Inf{} exceptions~\cite{github-issues}.

We introduce \FlowFPX{}, a toolkit for debugging
\fp{} exceptions in an easy and intuitive way~(\cref{s:flowfpx}):
\begin{itemize}
  \item
    The centerpiece of \FlowFPX{} is \FT{}, a dynamic analysis tool that uses
    Julia's operator overloading to selectively monitor for exceptions and to
    fuzz code for vulnerabilities.
  \item
    To visualize results, \FT{} adapts coalesced stack-trace graphs 
    (CSTGs, or \CSTG{}s)~\cite{humphreySystematicDebuggingMethods2014} for a whole-program view on the generation,
    propagation, and death of exceptional values.
  \item A companion tool \GPUFPX{}~\cite{llsflg-hpdc-2023} provides fine-grained insights for GPU kernels.
\end{itemize}
%
Together, the \FlowFPX{} tools have helped debug exceptions in a variety
of applications: from ocean simulations to heat models~(\cref{s:casestudies}).
That said, there is much more to explore.
\FlowFPX{} is a growing project and we welcome feedback from domain experts
on how to best suit their workflow.

This paper concludes with related work~(\cref{s:related}) and a
brief discussion~(\cref{s:discussion}).

@@

\subsection{Enhanced Fuzzing}

@@

>>>>

\begin{abstract}
  Reliable numerical computations are central to high-performance computing and machine learning.
  However, floating-point numbers exhibit non-intuitive behaviors that make writing correct numerical computation difficult.
  In particular, \emph{floating-point exceptions} occur when no mathematically sensible result is available, and it is difficult to find \emph{where} these exceptions occur.
  Unhandled exceptions can cause programs crash, exhibit bad behavior, or worse, silently produce good-looking but incorrect data.
  We present FlowFPX: a tool set for debugging floating-point code.
  FlowFPX can track floating-point exceptions occurring in Julia code to their sources.
  FlowFPX also detects otherwise silent exceptions which can foul program outputs.
  FlowFPX produces intuitive visualizations of summarized exception flows including how they are generated, propagated and killed, thus helping with debugging.
\end{abstract}

@@

Reliable numeric computations are central to high-performance computing,
machine learning, and scientific applications.
Yet, the \fp{} arithmetic that powers these computations is fundamentally
unreliable~(\cref{s:background}).
Exceptional values, such as Not a Number (\Nan{}) and infinity (\Inf{}),
can and often do arise thanks to culprits such as roundoff error,
catastrophic cancellation, singular matrices, and vanishing
derivatives~\cite{sdjmrstp-pc-2022,ddghlllprr-correctness-2022,gllprt-correctness-2021,fpchecker-reports,llg-soap-2022,bllmg-xloop-2022}.
When exceptions happen, domain experts must take time away from their main
focus and trek down a rabbit hole to find the source of the error.
There is little tool support to assist in this difficult task,
thus (unsurprisingly) a quick GitHub search reports over 4,000 open issues
related to \NaN{} or \Inf{} exceptions~\cite{github-issues}.

We introduce \FlowFPX{}, a toolkit for debugging
\fp{} exceptions~(\cref{s:flowfpx}):
\begin{itemize}
  \item
    The centerpiece of \FlowFPX{} is \FT{}, a dynamic analysis tool that uses
    Julia's operator overloading to selectively monitor for exceptions and to
    fuzz code for vulnerabilities.
  \item
    To visualize results, \FT{} adapts coalesced stack-trace graphs 
    (CSTGs, or \CSTG{}s)~\cite{humphreySystematicDebuggingMethods2014} for a whole-program view on the generation,
    propagation, and death of exceptional values.
  \item A companion tool \GPUFPX{}~\cite{llsflg-hpdc-2023} provides fine-grained insights for GPU kernels.
\end{itemize}
%
\FlowFPX{} has helped debug exceptions in a variety
of applications: from ocean simulations to heat models~(\cref{s:casestudies}).
This paper concludes with related work~(\cref{s:related}) and a
brief discussion~(\cref{s:discussion}).

@@

\subsection{Enhanced fuzzing}

@@

\subsection{Future work}

\FlowFPX{} is a growing project and we welcome feedback from domain experts
on how to best suit their workflow.
If you try any of our tools, please send us comments, suggestions, and bug reports via the tools' repositories on GitHub.
